---
title: "Week 2 Project"
author: "Javier González Onrubia"
date: "Apr. 5th 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## R Markdown and Leaflet

#### Dataset information
This project uses the [_Street Trees of the city of Barcelona_](http://opendata-ajuntament.barcelona.cat/data/en/dataset/arbrat-viari) 
dataset from the [Barcelona Open Data service](http://opendata-ajuntament.barcelona.cat/en/)
and is used under the [Creative Commons Attribution 4.0 license](https://creativecommons.org/licenses/by/4.0/).


```{r bodtrees_download, cache = TRUE}
df <- read.csv(url("http://opendata-ajuntament.barcelona.cat/resources/bcn/arbrat_viari.csv"))
names(df)
```

The dataset includes all trees that do not belong to parks or green
areas and is composed of `r nrow(df)` trees with `r ncol(df)`
variables each. 

We'll be using the following variables for this project:

* 'LATITUD_WGS84': Latitude in [WGS84 format](https://en.wikipedia.org/wiki/WGS84)
* 'LONGITUD_WGS84': Longitude in [WGS84 format](https://en.wikipedia.org/wiki/WGS84)
* 'NOM_CIENTIFIC': Tree's scientific name. This field is used on the
marker popup and alse to generate the link to the corresponding
Wikipedia article when clicked.
* 'ALCADA': A factor indicating the tree's height or _EXEMPLAR_ if
it's an illustratory tree for the species.

As the total number of trees will generate a too big HTML page, we
use the `runif()` function to pick 1000 trees to plot.

```{r bodtrees_datacleaning}
library(dplyr)
df <- select(df, 
             LATITUD_WGS84, LONGITUD_WGS84, NOM_CIENTIFIC, ALCADA)
head(unique(df$NOM_CIENTIFIC))
set.seed(12345)
df <- df[runif(1000,min = 1, max = nrow(df)),]
df$ALCADA <- relevel(df$ALCADA, 
                     "EXEMPLAR", "GRAN", "MITJANA", "PETITA")
```

### Leaflet map
We use the 'addAwesomeMarkers' function to change color and icon
of the marker as it provides a very convenient icon library
instead of downloading it. Please see [the Awesome Markers page on github](https://github.com/lvoogdt/Leaflet.awesome-markers) and 
[Ion icons](http://ionicons.com/) and 

```{r leaflet_bodtrees}
library(leaflet)

treesPopup <- paste(df$NOM_CIENTIFIC,
                    " <a href='http://en.wikipedia.org/wiki/",
                    df$NOM_CIENTIFIC,"'>[Wikipedia]</a>")
treecolors <- c("gold", "chartreuse", "seagreen", "darkgreen")
        
getTreeColor <- function(heights) {
    sapply(heights, function(h) {
        if(is.na(h) | is.null(h)) {"white"}
        else treecolors[h]
    })
}

icons <- awesomeIcons(
    icon = 'leaf',
    iconColor = 'black',
    library = 'ion',
    markerColor = getTreeColor(df$ALCADA)
)

df %>% leaflet() %>%
       addTiles() %>%
       addAwesomeMarkers(lat = df$LATITUD_WGS84, 
                         lng = df$LONGITUD_WGS84,
                         popup = treesPopup,
                         icon = icons,
                         clusterOptions = markerClusterOptions()) %>%
       addLegend(title = "Tree Height legend",
                 labels = c("Exemplar", "Tall", "Medium", "Small"),
                 colors = treecolors)
```


